### 2023-05-20 DMTN-065.tex
Update DMTN-065.tex

\documentclass[DM,lsstdoc,toc]{lsstdoc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{latexsym}
\usepackage{enumitem}

\title[LSST Special Programs]{Data Management for LSST Special Programs}

\author{M.~L.~Graham, M.~Juri\'{c}, K.-T.~Lim, E.~Bellm, G.~Dubois-Felsmann, L.~P.~Guy, and the Data Management System Science Team}

\setDocRef{DMTN-065}
\date{\today}
\setDocUpstreamLocation{\url{https://github.com/lsst-dm/dmtn-065}}

\setDocAbstract{
This document provides a summary of the requirements on Rubin Data Management for processing and serving data products
related to Special Programs: additional survey areas and/or observing strategies that are driven by specific science 
goals which build on, or are beyond, the core science pillars of the LSST.

Recommendations for the specific implementations needed to fulfill these requirements and maximize science from
Special Programs are provided on a case-by-case basis, with examples.

Hardware and processing boundaries on the potential diversity of data from Special Programs 
(e.g., longer or shorter exposure times), and the ability of the Data Management System to
process this data, and also provided along with a discussion of when user-generated processing 
and data products will be needed to meet some science goals.

The main target audience of this document is Rubin Observatory staff -- the construction- and operations-era Data Management
teams -- but members of the science community who are planning to use Special Programs to reach their science goals may also 
find this document useful.
}

\setDocChangeRecord{%
\addtohist{0}{2017-11-14}{Status: internal working document.}{Melissa Graham}
\addtohist{1}{2018-06-17}{Updated to finalize and issue.}{Melissa Graham}
\addtohist{2}{2021-12-01}{Updates per DM-20375.}{Melissa Graham}
\addtohist{3}{2023-05-31}{Updates per DM-32723.}{Melissa Graham}
}

\begin{document}

\maketitle

% CITATION EXAMPLES
% \verb|\citellp|: \citellp{LPM-17, LSE-30} \\
% \verb|\citell|: (SRD; \citell{LPM-17,LSE-29}) \\
% \verb|\citep[][]|: \citep[e.g.,][are interesting]{LPM-17,LSE-29} \\
% \verb|\cite|: \cite{LPM-17,LSE-29}

\input{sec_intro.tex}
x
\input{sec_sci.tex}

\input{sec_proc.tex}

\input{sec_req.tex}

% % % % % % % % % % % % % % % % % %
\clearpage
\bibliography{local,lsst,refs,books,refs_ads}

% % % % % % % % % % % % % % % % % %
\clearpage
\appendix

\input{ap_spcs.tex}

\input{ap_procbounds.tex}

\input{ap_hardbounds.tex}

\input{ap_docrev.tex}

\end{document}