### 2023-05-20 ap_spcs.tex
Update ap_spcs.tex

\section{Special Programs Processing Examples}\label{sec:spcs}

Step-by-step examples of the processing, either by RDM or users, 
for data from several Special Programs.

Basic steps that we use to describe a processing case study: \\
Step 1. Data acquisition. \\
Step 2. RDM prompt processing and alert production. \\
Step 3. RDM special processing with reconfigured pipelines. \\
Step 4. RDM inclusion in the WFD Data Release data products. \\
Step 5. User processing and user-generated data products. \\

\subsection{Searching for TNOs with Shift-and-Stack}\label{ssec:SPCS_TNO}

This Special Programs processing summary is based on Becker et al. (2011) 
white paper to find TNOs with shift-and stack (SAS) \citedsp{Document-11013}.

Step 1. Data acquisition. \\
The observational sequence is triggered. 
In a single night, the 9 adjacent fields in a 3x3 grid are observed with 
$336$ $\times$ $15$ second $r$-band exposures. 
This sequence is always repeated 2-3 nights later. 
This re-visit sequence is repeated 3 more times: 1.5 months, 3 months, and 13.5 months later. 
Data obtained in the $g$-band filter is also acceptable.

Step 2. RDM prompt processing and alert production. \\
Each $2\times15$ second standard visit is processed by the Prompt pipeline 
and alerts are released within 60 seconds.
Within 24 hours, the {\tt DiaSource} and {\tt DiaObject} catalogs are updated
to include the results of Prompt processing of these visits.
After 80 hours, the processed visit images and difference images become available.
All images and sources originating from this Special Program have 
region and program labels, e.g., ``TNO Survey".

Note that the results of prompt processing is not very relevant for this Special Program,
which requires a year of dispersed observations before the processing pipelines 
for shift-and-stack can be run, but including them in prompt processing means
they can contribute to LSST's other time-domain and Solar System science goals.

Step 3. RDM special processing with reconfigured pipelines. \\
None possible. 
Shift-and-stack processing is beyond the scope of DM's algorithms.

Step 4. RDM inclusion in the WFD Data Release data products. \\
All standard visits are reprocessed and included in the annual Data Release.
Data products would include processed visit images and {\tt Source} and {\tt ForcedSource}
catalog rows, along with difference images and {\tt DiaSource} catalog rows.
All images and sources originating from this Special Program have 
region and program labels, e.g., ``TNO Survey".
As with all Special Programs data, these visits might also be included in the 
WFD deeply coadded images if RDM decides it is both possible and beneficial.

Step 5. User processing and user-generated data products. \\
The user-generated pipeline running the shift-and-stack processing would be set up and submitted 
for batch processing by the user through the Science Platform or on an external processor. 
Pipeline inputs would be the 336 processed exposures per field per re-visit sequence. 
The LSST Science Pipelines difference imaging routine could be used with the same template tract/patch for all. 
User-generated algorithms will shift the exposures and create difference images, and then the LSST
Science Pipelines could be used to stack and do source detection and characterization, 
and generate an object database. 
Custom code will derive orbital parameters for the detections (or perhaps interface with the MPC),
and store them in a separate {\tt SSObjects}-like database.

\subsection{Searching for Supernovae in Deep Drilling Fields}\label{ssec:SPCS_SNDDF}

Step 1. Data acquisition. \\
On a single deep drilling field, the scheduler obtains e.g., 5, 10, 10, 9, and 10 standard visits 
with $2\times15$ second exposures in $grizy$ (or similar for the night's filter set) 
and a small dither pattern between visits.

Step 2. RDM prompt processing and alert production. \\
As in the example above, but all images and sources originating from this Special Program would have 
region and program labels of, e.g., ``DDF Name".

Step 3. RDM special processing with reconfigured pipelines. \\
The required data products for this science goal can be met by reconfiguring the DM pipelines. 
First, a template image of appropriate depth for ``nightly" difference imaging 
would be made using RDM stacking algorithms. 
On nights when this DDF is observed, at the end of the sequence of observations, 
reconfigured RDM algorithms would be automatically triggered to create a nightly deep stack, 
PSF-match it with the appropriately-made template, 
create a nightly deep difference image, run source detection on it, and update separate 
{\tt DiaObject}- and {\tt DiaSource}-like catalogs.
All of this could happen in, e.g., 24 hours, but note there is no requirement on this timescale.
On an annual basis, deeply coadded images that include all of the DDF's visits to dates would
be generated, along with a separate {\tt Object}-like catalog for the DDF.

Step 4. RDM inclusion in the WFD Data Release data products. \\
As in the example above, but all images and sources originating from this Special Program would have 
region and program labels of, e.g., ``DDF Name".

Step 5. User processing and user-generated data products. \\
For the science goal of searching for supernovae in nightly stacked DDF images, 
no separate user-generated processing or data products appear to be necessary.

\subsection{A Twilight Survey with Short Exposures}\label{ssec:SPCS_Twilight}

Several kinds of twilight surveys with short exposures have been or might be proposed, 
to put brighter stars (or transients such as supernovae) that saturate in a $15$ second image onto the 
LSST photometric system and/or to take observations that are particularly well suited 
(e.g., are 60 degrees from the sun) for finding Near-Earth Objects (NEOs).
The processing case study for these is currently limited by unknowns.

Step 1. Data acquisition. \\
At a specified time (or e.g., 6 degree twilight), the scheduler begins dither pattern of short exposures. 
Location and exposure times are set by the sky brightness and desired saturation limits.

Step 2. RDM prompt processing and alert production. \\
Pending studies of short-exposure suitability for DIA (see Section \ref{sec:procbounds}) 
these data could {\it potentially} be processed by the Prompt pipeline and lead to alerts.

Step 3. RDM special processing with reconfigured pipelines. \\
So long as the short-exposure images can be processed and have enough stars for photometric 
and astrometric calibration, reconfigured DM pipelines will probably be sufficient for 
creating image and catalog products from this kind of data.

Step 4. RDM inclusion in the WFD Data Release data products. \\
These short-exposure, high sky background images would not contribute to the DRP data products created for the WFD survey.

Step 5. User processing and user-generated data products. \\
If short-exposure images cannot be processed with the existing DM algorithms, 
user-generated processing would be needed to reduce the raw data.

\textbf{Side note:} A short-exposure survey of the bright stars of M67, described in Chapter 10.4 of the 
Observing Strategy White Paper \citep{2017arXiv170804058L}, suggests using the stretch goal of 
0.1 second exposures or, if that is not possible, \textit{"custom pixel masks to accurately perform 
photometry on stars as much as 6 magnitudes brighter than the saturation level"}. 
This would need user-generated processing.