### 2023-10-16 DMTN-065.tex
shorten abstract

\documentclass[DM,lsstdoc,toc]{lsstdoc}
\usepackage{graphicx}
\usepackage{url}
\usepackage{latexsym}
\usepackage{enumitem}

\title[LSST Special Programs]{Data Management for LSST Special Programs}

\author{M.~L.~Graham, M.~Juri\'{c}, K.-T.~Lim, E.~Bellm, G.~Dubois-Felsmann, L.~P.~Guy, and the Data Management System Science Team}

\setDocRef{DMTN-065}
\date{\today}
\setDocUpstreamLocation{\url{https://github.com/lsst-dm/dmtn-065}}

\setDocAbstract{
Special Programs are additional survey areas and/or observing strategies that are driven by specific science 
goals which build on, or are beyond, the core science pillars of the LSST.
This document provides a summary of the requirements on Rubin Data Management for processing and serving data products
related to Special Programs.

\smallskip
Recommendations for the specific implementations needed to fulfill these requirements and maximize science from
Special Programs are provided on a case-by-case basis, with examples.
Hardware and processing boundaries on the potential diversity of data from Special Programs
are also provided, along with a discussion of when user-generated processing 
and data products will be needed to meet Special Programs' science goals.
}

\setDocChangeRecord{%
\addtohist{0}{2017-11-14}{Status: internal working document.}{Melissa Graham}
\addtohist{1}{2018-06-17}{Updated to finalize and issue.}{Melissa Graham}
\addtohist{2}{2021-12-01}{Updates per DM-20375.}{Melissa Graham}
\addtohist{3}{2023-10-31}{Updates per DM-32723.}{Melissa Graham}
}

\begin{document}

\maketitle

% CITATION EXAMPLES
% \verb|\citellp|: \citellp{LPM-17, LSE-30} \\
% \verb|\citell|: (SRD; \citell{LPM-17,LSE-29}) \\
% \verb|\citep[][]|: \citep[e.g.,][are interesting]{LPM-17,LSE-29} \\
% \verb|\cite|: \cite{LPM-17,LSE-29}

\input{sec_intro.tex}

\input{sec_sci.tex}

\input{sec_proc.tex}

\input{sec_req.tex}

% % % % % % % % % % % % % % % % % %
\bibliography{local,lsst,refs,books,refs_ads}

% % % % % % % % % % % % % % % % % %
\appendix

\input{ap_spcs.tex}

\input{ap_procbounds.tex}

\input{ap_hardbounds.tex}

\input{ap_docrev.tex}

\end{document}